### 2015-01-03 mcmc.tex
tweaks
importance, nested, and f() is posterior

% to-do
% -----
% - where to put comments about how to structure code ln_likelihood(pars, data, I) and ln_prior(pars, I)
% - need to mention case where there is an MC integral INSIDE the likelihood function
% - need to mention testing the likelihood function
% - send to friendlies for comments
%   - Lang
%   - Rix
%   - Bovy
%   - Brewer
%   - Hill
% - post on arXiv

% style notes
% -----------
% - [LaTeX] newline after every full stop for proper git diffing
% - [LaTeX] \eqnarray not \equation
% - pdf not PDF
% - is it ``a MCMC sampling'' or ``an MCMC sampling''?
% - \note{} right after word if you are endnoting a phrase,
%   but after the full stop if endnoting the sentence or paragraph.
% - do we work in $x$ or $\pars$?  I am confused...
% - should we use the word ``walker'' anywhere, everywhere, or never?

@@

Finally---and perhaps most importantly---a magical property of a sampling
  (in a $d$-dimensional space)
  is that if some of your $d$ dimensions are totally uninteresting
  (nuisance parameters, if you will)
  and some of your $d$ dimensions are of great interest,
  the sampling in the full $d$-space is trivially converted into a sampling in the subspace of interest:
You just drop from each $x_k$ vector (or blob or list) the dimensions of no interest!
That is, the projection of the sampling to the subspace of interest produces
  a sampling of the marginalized pdf, marginalizing (or projecting) out the nuisance parameters.%
  \note{This point about marginalization is, once again, a use of the sampling to perform an \emph{integral};
        the marginalized pdf is obtained from the full pdf by an integration.
        The sampling performs this integration automatically.}
That is extremely important for inference,
  where there are always parameters with very different levels of importance to the scientific conclusions.
This point will return again below, either explicitly or implicitly. % does it?

@@

\paragraph{Importance:}
... HOGG

@@

\paragraph{Nested:}
... HOGG

@@

For a committed probabilistic scientist---frequentist or
bayesian---there is no ``answer'', there are only probability density
functions (pdfs).  Any time in a paper or in an email or in a conversation
we say what ``the answer is''---even if we say it with an error
bar---we have departed the probabilistic program.  There are many
principled ways to depart the program; in another \documentname, we
discuss the economic explanation of how we can make hard decisions in
the context of probabilistic reasoning\note{CITE DECISION DOCUMENT}.
But without going into that ``economic'' model, it is fair to say that a
substantial problem with being a committed probabilist is that when we
\emph{publish}, we are not permitted (not now, at least) to publish a
\emph{probability distribution over pubications}; we have to publish
\emph{one single, deterministic text}; we have to make a decision about what
to write in the title, the abstract, the tables, figures, and results
section.  Given a MCMC sampling of the posterior pdf, what do we
report as our \emph{results}?

>>>>

% to-do
% -----
% - perhaps re-order to m-h, convergence, reporting results, other methods ??
% - where to put comments about how to structure code ln_likelihood(pars, data, I) and ln_prior(pars, I)
% - need to mention case where there is an MC integral INSIDE the likelihood function
% - need to mention testing the likelihood function
% - send to friendlies for comments
%   - Lang
%   - Rix
%   - Bovy
%   - Brewer
%   - Hill
% - post on arXiv

% style notes
% -----------
% - [LaTeX] newline after every full stop for proper git diffing
% - [LaTeX] eqnarray not equation
% - pdf not PDF
% - is it ``a MCMC sampling'' or ``an MCMC sampling''?
% - \note{} right after word if you are endnoting a phrase,
%   but after the full stop if endnoting the sentence or paragraph.
% - do we work in $x$ or $\pars$?  I am confused...
% - should we use the word ``walker'' anywhere, everywhere, or never?

@@

Finally---and perhaps most importantly---a magical property of a sampling
  (in a $d$-dimensional space)
  is that if some of your $d$ dimensions are totally uninteresting
  (nuisance parameters, if you will)
  and some of your $d$ dimensions are of great interest,
  the sampling in the full $d$-space is trivially converted into a sampling in the subspace of interest:
You just drop from each $x_k$ vector (or blob or list) the dimensions of no interest!
That is, the projection of the sampling to the subspace of interest produces
  a sampling of the marginalized pdf, marginalizing (or projecting) out the nuisance parameters.%
  \note{This point about marginalization is, once again, a use of the sampling to perform an \emph{integral};
        the marginalized pdf is obtained from the full pdf by an integration.
        The sampling performs this integration automatically.}
That is extremely important for inference,
  where there are always parameters with very different levels of importance to the scientific conclusions.
This point will return again below, either explicitly or implicitly. % HOGG: does it?

Although the discussion in this \documentname\ is general, the most
common use of MCMC sampling (for us, anyway) is in probabilistic
inference.
For this reason, we will often refer to the function $f(\pars)$
colloquially as ``the posterior pdf''\note{We will also sometimes
  refer to expectations under $f(\pars)$ as posterior means, and
  medians as median-of-posterior values, and so on.} even though it is
implicitly ill-normalized and might not be a posterior pdf in any
sense.
We will also occasionally assume---just because it is true in
inference---that the function $f(\theta)$ is the product of two
functions, one called ``the prior pdf'' and one called ``the
likelihood''.
Again, this usage is colloquial and is only strictly correct in
inference contexts with proper inputs.
That the function $f(\theta)$ can be thought of as a product of a
prior pdf and a likelihood is only necessary for what follows in
the context of advanced sampling techniques like tempering or
nested sampling, both mentioned briefly below.

@@

\paragraph{Slice:}
... DFM, can you help here?

\paragraph{Importance:}
Sometimes it is possible to sample efficiently from a distribution
that is close to the target function $f(\pars)$ you care about but
hard to sample efficiently with MCMC.
An example is a problem that has a simple (Gaussian, say) likelihood
function times a non-trivial prior:
It is easy to sample from a Gaussian likelihood times a Gaussian
prior, even without MCMC, but it might be hard to do the same sampling
under a more complex prior.
Importance sampling is a direct sampling method (not really an MCMC
method at all):
You take a sampling from the approximate function (which by assumption
was easy) and then you re-weight the sampling (or do a subsequent
rejection of samples) using the ratio of the true function to the
approximate function as a weight or probability.
We use this quite a bit in our hierarchical inferences.\note{HOGG:
  Citations}

The only significant technical details for importance sampling are the
following:
The approximate function must have the same (or greater) support as
the true function, and the method degrades in efficiency as the two
functions diverge.
If the ratio of the true function to the approximate function is to be
used as a probability for rejection of samples, then the true function
must be greater than or equal to the approximate function everywhere!

@@

\paragraph{Nested:}
Similar to tempering are a class of methods called ``nested''
samplings, which also smoothly change the target distribution.
In the nested case, instead of increasing the temperature, the
samplings are of a censored version of the prior, censored by the
value of the likelihoood function.

@@

For a committed probabilistic scientist---frequentist or
Bayesian---there is no ``answer'', there are only probability density
functions (pdfs).  Any time in a paper or in an email or in a conversation
we say what ``the answer is''---even if we say it with an error
bar---we have departed the probabilistic program.  There are many
principled ways to depart the program; in another \documentname, we
discuss the economic explanation of how we can make hard decisions in
the context of probabilistic reasoning\note{CITE DECISION DOCUMENT}.
But without going into that ``economic'' model, it is fair to say that a
substantial problem with being a committed probabilist is that when we
\emph{publish}, we are not permitted (not now, at least) to publish a
\emph{probability distribution over pubications}; we have to publish
\emph{one single, deterministic text}; we have to make a decision about what
to write in the title, the abstract, the tables, figures, and results
section.  Given a MCMC sampling of the posterior pdf, what do we
report as our \emph{results}?
